### 2021-12-11 PaperForReview.tex
initial commit

% CVPR 2022 Paper Template
% based on the CVPR template provided by Ming-Ming Cheng (https://github.com/MCG-NKU/CVPR_Template)
% modified and extended by Stefan Roth (stefan.roth@NOSPAMtu-darmstadt.de)

\documentclass[10pt,twocolumn,letterpaper]{article}

%%%%%%%%% PAPER TYPE  - PLEASE UPDATE FOR FINAL VERSION
\usepackage[review]{cvpr}      % To produce the REVIEW version
%\usepackage{cvpr}              % To produce the CAMERA-READY version
%\usepackage[pagenumbers]{cvpr} % To force page numbers, e.g. for an arXiv version

% Include other packages here, before hyperref.
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}

% It is strongly recommended to use hyperref, especially for the review version.
% hyperref with option pagebackref eases the reviewers' job.
% Please disable hyperref *only* if you encounter grave issues, e.g. with the
% file validation for the camera-ready version.
%
% If you comment hyperref and then uncomment it, you should delete
% ReviewTempalte.aux before re-running LaTeX.
% (Or just hit 'q' on the first LaTeX run, let it finish, and you
%  should be clear).
\usepackage[pagebackref,breaklinks,colorlinks]{hyperref}

% Support for easy cross-referencing
\usepackage[capitalize]{cleveref}
\crefname{section}{Sec.}{Secs.}
\Crefname{section}{Section}{Sections}
\Crefname{table}{Table}{Tables}
\crefname{table}{Tab.}{Tabs.}

%%%%%%%%% PAPER ID  - PLEASE UPDATE
\def\cvprPaperID{*****} % *** Enter the CVPR Paper ID here
\def\confName{ECE 251C}
\def\confYear{Fall 2021}

\begin{document}

%%%%%%%%% TITLE - PLEASE UPDATE
\title{Feature Extraction of ECG signals using Wavelet transform \confName~Final Project report}

\author{Hari Prasad Sankar\\
UCSD\\
Institution1 address\\
{\tt\small hsankar@ucsd.edu}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
\and
Abhishek Khare\\
UCSD\\
First line of institution2 address\\
{\tt\small akhare@ucsd.edu}
}
\maketitle

%%%%%%%%% ABSTRACT
\begin{abstract}
   The ABSTRACT is to be in fully justified italicized text, at the top of the left-hand column, below the author and affiliation information.
   Use the word ``Abstract'' as the title, in 12-point Times, boldface type, centered relative to the column, initially capitalized.
   The abstract is to be in 10-point, single-spaced type.
   Leave two blank lines after the Abstract, then begin the main text.
   Look at previous CVPR abstracts to get a feel for style and length.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}
\label{sec:intro}

Please follow the steps outlined below when submitting your manuscript to the IEEE Computer Society Press.
This style guide now has several important modifications (for example, you are no longer warned against the use of sticky tape to attach your artwork to the paper), so all authors should read this new version.

%-------------------------------------------------------------------------
\subsection{Language}

All manuscripts must be in English.

\subsection{Dual submission}

Please refer to the author guidelines on the \confName\ \confYear\ web page for a
discussion of the policy on dual submissions.

\subsection{Paper length}
Papers, excluding the references section, must be no longer than eight pages in length.
The references section will not be included in the page count, and there is no limit on the length of the references section.
For example, a paper of eight pages with two pages of references would have a total length of 10 pages.
{\bf There will be no extra page charges for \confName\ \confYear.}

Overlength papers will simply not be reviewed.
This includes papers where the margins and formatting are deemed to have been significantly altered from those laid down by this style guide.
Note that this \LaTeX\ guide already sets figure captions and references in a smaller font.
The reason such papers will not be reviewed is that there is no provision for supervised revisions of manuscripts.
The reviewing process cannot determine the suitability of the paper for presentation in eight pages if it is reviewed in eleven.

%-------------------------------------------------------------------------
\subsection{The ruler}
The \LaTeX\ style defines a printed ruler which should be present in the version submitted for review.
The ruler is provided in order that reviewers may comment on particular lines in the paper without circumlocution.
If you are preparing a document using a non-\LaTeX\ document preparation system, please arrange for an equivalent ruler to appear on the final output pages.
The presence or absence of the ruler should not change the appearance of any other content on the page.
The camera-ready copy should not contain a ruler.
(\LaTeX\ users may use options of cvpr.sty to switch between different versions.)

Reviewers:
note that the ruler measurements do not align well with lines in the paper --- this turns out to be very difficult to do well when the paper contains many figures and equations, and, when done, looks ugly.
Just use fractional references (\eg, this line is $087.5$), although in most cases one would expect that the approximate location will be adequate.

\subsection{Paper ID}
Make sure that the Paper ID from the submission system is visible in the version submitted for review (replacing the ``*****'' you see in this document).
If you are using the \LaTeX\ template, \textbf{make sure to update paper ID in the appropriate place in the tex file}.

\subsection{Mathematics}

Please number all of your sections and displayed equations as in these examples:
\begin{equation}
  E = m\cdot c^2
  \label{eq:important}
\end{equation}
and
\begin{equation}
  v = a\cdot t.
  \label{eq:also-important}
\end{equation}
It is important for readers to be able to refer to any particular equation.
Just because you did not refer to it in the text does not mean some future reader might not need to refer to it.
It is cumbersome to have to use circumlocutions like ``the equation second from the top of page 3 column 1''.
(Note that the ruler will not be present in the final copy, so is not an alternative to equation numbers).
All authors will benefit from reading Mermin's description of how to write mathematics:
\url{http://www.pamitc.org/documents/mermin.pdf}.

\subsection{Blind review}

Many authors misunderstand the concept of anonymizing for blind review.
Blind review does not mean that one must remove citations to one's own work---in fact it is often impossible to review a paper unless the previous citations are known and available.

Blind review means that you do not use the words ``my'' or ``our'' when citing previous work.
That is all.
(But see below for tech reports.)

Saying ``this builds on the work of Lucy Smith [1]'' does not say that you are Lucy Smith;
it says that you are building on her work.
If you are Smith and Jones, do not say ``as we show in [7]'', say ``as Smith and Jones show in [7]'' and at the end of the paper, include reference 7 as you would any other cited work.

An example of a bad paper just asking to be rejected:
\begin{quote}
\begin{center}
    An analysis of the frobnicatable foo filter.
\end{center}

   In this paper we present a performance analysis of our previous paper [1], and show it to be inferior to all previously known methods.
   Why the previous paper was accepted without this analysis is beyond me.

   [1] Removed for blind review
\end{quote}

An example of an acceptable paper:
\begin{quote}
\begin{center}
     An analysis of the frobnicatable foo filter.
\end{center}

   In this paper we present a performance analysis of the  paper of Smith \etal [1], and show it to be inferior to all previously known methods.
   Why the previous paper was accepted without this analysis is beyond me.

   [1] Smith, L and Jones, C. ``The frobnicatable foo filter, a fundamental contribution to human knowledge''. Nature 381(12), 1-213.
\end{quote}

If you are making a submission to another conference at the same time, which covers similar or overlapping material, you may need to refer to that submission in order to explain the differences, just as you would if you had previously published related work.
In such cases, include the anonymized parallel submission~\cite{Authors14} as supplemental material and cite it as
\begin{quote}
[1] Authors. ``The frobnicatable foo filter'', F\&G 2014 Submission ID 324, Supplied as supplemental material {\tt fg324.pdf}.
\end{quote}

Finally, you may feel you need to tell the reader that more details can be found elsewhere, and refer them to a technical report.
For conference submissions, the paper must stand on its own, and not {\em require} the reviewer to go to a tech report for further details.
Thus, you may say in the body of the paper ``further details may be found in~\cite{Authors14b}''.
Then submit the tech report as supplemental material.
Again, you may not assume the reviewers will read this material.

Sometimes your paper is about a problem which you tested using a tool that is widely known to be restricted to a single institution.
For example, let's say it's 1969, you have solved a key problem on the Apollo lander, and you believe that the CVPR70 audience would like to hear about your
solution.
The work is a development of your celebrated 1968 paper entitled ``Zero-g frobnication: How being the only people in the world with access to the Apollo lander source code makes us a wow at parties'', by Zeus \etal.

You can handle this paper like any other.
Do not write ``We show how to improve our previous work [Anonymous, 1968].
This time we tested the algorithm on a lunar lander [name of lander removed for blind review]''.
That would be silly, and would immediately identify the authors.
Instead write the following:
\begin{quotation}
\noindent
   We describe a system for zero-g frobnication.
   This system is new because it handles the following cases:
   A, B.  Previous systems [Zeus et al. 1968] did not  handle case B properly.
   Ours handles it by including a foo term in the bar integral.

   ...

   The proposed system was integrated with the Apollo lunar lander, and went all the way to the moon, don't you know.
   It displayed the following behaviours, which show how well we solved cases A and B: ...
\end{quotation}
As you can see, the above text follows standard scientific convention, reads better than the first version, and does not explicitly name you as the authors.
A reviewer might think it likely that the new paper was written by Zeus \etal, but cannot make any decision based on that guess.
He or she would have to be sure that no other authors could have been contracted to solve problem B.
\medskip

\noindent
FAQ\medskip\\
{\bf Q:} Are acknowledgements OK?\\
{\bf A:} No.  Leave them for the final copy.\medskip\\
{\bf Q:} How do I cite my results reported in open challenges?
{\bf A:} To conform with the double-blind review policy, you can report results of other challenge participants together with your results in your paper.
For your results, however, you should not identify yourself and should not mention your participation in the challenge.
Instead present your results referring to the method proposed in your paper and draw conclusions based on the experimental comparison to other results.\medskip\\

\begin{figure}[t]
  \centering
  \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
   %\includegraphics[width=0.8\linewidth]{egfigure.eps}

   \caption{Example of caption.
   It is set in Roman so that mathematics (always set in Roman: $B \sin A = A \sin B$) may be included without an ugly clash.}
   \label{fig:onecol}
\end{figure}

\subsection{Miscellaneous}

\noindent
Compare the following:\\
\begin{tabular}{ll}
 \verb'$conf_a$' &  $conf_a$ \\
 \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
\end{tabular}\\
See The \TeX book, p165.

The space after \eg, meaning ``for example'', should not be a sentence-ending space.
So \eg is correct, {\em e.g.} is not.
The provided \verb'\eg' macro takes care of this.

When citing a multi-author paper, you may save space by using ``et alia'', shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}'' is a complete word).
If you use the \verb'\etal' macro provided, then you need not worry about double periods when used at the end of a sentence as in Alpher \etal.
However, use it only when there are three or more authors.
Thus, the following is correct:
   ``Frobnication has been trendy lately.
   It was introduced by Alpher~\cite{Alpher02}, and subsequently developed by
   Alpher and Fotheringham-Smythe~\cite{Alpher03}, and Alpher \etal~\cite{Alpher04}.''

This is incorrect: ``... subsequently developed by Alpher \etal~\cite{Alpher03} ...'' because reference~\cite{Alpher03} has just two authors.

% Update the cvpr.cls to do the following automatically.
% For this citation style, keep multiple citations in numerical (not
% chronological) order, so prefer \cite{Alpher03,Alpher02,Authors14} to
% \cite{Alpher02,Alpher03,Authors14}.

\begin{figure*}
  \centering
  \begin{subfigure}{0.68\linewidth}
    \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
    \caption{An example of a subfigure.}
    \label{fig:short-a}
  \end{subfigure}
  \hfill
  \begin{subfigure}{0.28\linewidth}
    \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
    \caption{Another example of a subfigure.}
    \label{fig:short-b}
  \end{subfigure}
  \caption{Example of a short caption, which should be centered.}
  \label{fig:short}
\end{figure*}

%------------------------------------------------------------------------
\section{Formatting your paper}
\label{sec:formatting}

All text must be in a two-column format.
The total allowable size of the text area is $6\frac78$ inches (17.46 cm) wide by $8\frac78$ inches (22.54 cm) high.
Columns are to be $3\frac14$ inches (8.25 cm) wide, with a $\frac{5}{16}$ inch (0.8 cm) space between them.
The main title (on the first page) should begin 1 inch (2.54 cm) from the top edge of the page.
The second and following pages should begin 1 inch (2.54 cm) from the top edge.
On all pages, the bottom margin should be $1\frac{1}{8}$ inches (2.86 cm) from the bottom edge of the page for $8.5 \times 11$-inch paper;
for A4 paper, approximately $1\frac{5}{8}$ inches (4.13 cm) from the bottom edge of the
page.

%-------------------------------------------------------------------------
\subsection{Margins and page numbering}

All printed material, including text, illustrations, and charts, must be kept
within a print area $6\frac{7}{8}$ inches (17.46 cm) wide by $8\frac{7}{8}$ inches (22.54 cm)
high.
%
Page numbers should be in the footer, centered and $\frac{3}{4}$ inches from the bottom of the page.
The review version should have page numbers, yet the final version submitted as camera ready should not show any page numbers.
The \LaTeX\ template takes care of this when used properly.

%-------------------------------------------------------------------------
\subsection{Type style and fonts}

Wherever Times is specified, Times Roman may also be used.
If neither is available on your word processor, please use the font closest in
appearance to Times to which you have access.

MAIN TITLE.
Center the title $1\frac{3}{8}$ inches (3.49 cm) from the top edge of the first page.
The title should be in Times 14-point, boldface type.
Capitalize the first letter of nouns, pronouns, verbs, adjectives, and adverbs;
do not capitalize articles, coordinate conjunctions, or prepositions (unless the title begins with such a word).
Leave two blank lines after the title.

AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title
and printed in Times 12-point, non-boldface type.
This information is to be followed by two blank lines.

The ABSTRACT and MAIN TEXT are to be in a two-column format.

MAIN TEXT.
Type main text in 10-point Times, single-spaced.
Do NOT use double-spacing.
All paragraphs should be indented 1 pica (approx.~$\frac{1}{6}$ inch or 0.422 cm).
Make sure your text is fully justified---that is, flush left and flush right.
Please do not place any additional blank lines between paragraphs.

Figure and table captions should be 9-point Roman type as in \cref{fig:onecol,fig:short}.
Short captions should be centred.

\noindent Callouts should be 9-point Helvetica, non-boldface type.
Initially capitalize only the first word of section titles and first-, second-, and third-order headings.

FIRST-ORDER HEADINGS.
(For example, {\large \bf 1. Introduction}) should be Times 12-point boldface, initially capitalized, flush left, with one blank line before, and one blank line after.

SECOND-ORDER HEADINGS.
(For example, { \bf 1.1. Database elements}) should be Times 11-point boldface, initially capitalized, flush left, with one blank line before, and one after.
If you require a third-order heading (we discourage it), use 10-point Times, boldface, initially capitalized, flush left, preceded by one blank line, followed by a period and your text on the same line.

%-------------------------------------------------------------------------
\subsection{Footnotes}

Please use footnotes\footnote{This is what a footnote looks like.
It often distracts the reader from the main flow of the argument.} sparingly.
Indeed, try to avoid footnotes altogether and include necessary peripheral observations in the text (within parentheses, if you prefer, as in this sentence).
If you wish to use a footnote, place it at the bottom of the column on the page on which it is referenced.
Use Times 8-point type, single-spaced.

%-------------------------------------------------------------------------
\subsection{Cross-references}

For the benefit of author(s) and readers, please use the
{\small\begin{verbatim}
  \cref{...}
\end{verbatim}}  command for cross-referencing to figures, tables, equations, or sections.
This will automatically insert the appropriate label alongside the cross-reference as in this example:
\begin{quotation}
  To see how our method outperforms previous work, please see \cref{fig:onecol} and \cref{tab:example}.
  It is also possible to refer to multiple targets as once, \eg~to \cref{fig:onecol,fig:short-a}.
  You may also return to \cref{sec:formatting} or look at \cref{eq:also-important}.
\end{quotation}
If you do not wish to abbreviate the label, for example at the beginning of the sentence, you can use the
{\small\begin{verbatim}
  \Cref{...}
\end{verbatim}}
command. Here is an example:
\begin{quotation}
  \Cref{fig:onecol} is also quite important.
\end{quotation}

%-------------------------------------------------------------------------
\subsection{References}

List and number all bibliographical references in 9-point Times, single-spaced, at the end of your paper.
When referenced in the text, enclose the citation number in square brackets, for
example~\cite{Authors14}.
Where appropriate, include page numbers and the name(s) of editors of referenced books.
When you cite multiple papers at once, please make sure that you cite them in numerical order like this \cite{Alpher02,Alpher03,Alpher05,Authors14b,Authors14}.
If you use the template as advised, this will be taken care of automatically.

\begin{table}
  \centering
  \begin{tabular}{@{}lc@{}}
    \toprule
    Method & Frobnability \\
    \midrule
    Theirs & Frumpy \\
    Yours & Frobbly \\
    Ours & Makes one's heart Frob\\
    \bottomrule
  \end{tabular}
  \caption{Results.   Ours is better.}
  \label{tab:example}
\end{table}

%-------------------------------------------------------------------------
\subsection{Illustrations, graphs, and photographs}

All graphics should be centered.
In \LaTeX, avoid using the \texttt{center} environment for this purpose, as this adds potentially unwanted whitespace.
Instead use
{\small\begin{verbatim}
  \centering
\end{verbatim}}
at the beginning of your figure.
Please ensure that any point you wish to make is resolvable in a printed copy of the paper.
Resize fonts in figures to match the font in the body text, and choose line widths that render effectively in print.
Readers (and reviewers), even of an electronic copy, may choose to print your paper in order to read it.
You cannot insist that they do otherwise, and therefore must not assume that they can zoom in to see tiny details on a graphic.

When placing figures in \LaTeX, it's almost always best to use \verb+\includegraphics+, and to specify the figure width as a multiple of the line width as in the example below
{\small\begin{verbatim}
   \usepackage{graphicx} ...
   \includegraphics[width=0.8\linewidth]
                   {myfile.pdf}
\end{verbatim}
}

%-------------------------------------------------------------------------
\subsection{Color}

Please refer to the author guidelines on the \confName\ \confYear\ web page for a discussion of the use of color in your document.

If you use color in your plots, please keep in mind that a significant subset of reviewers and readers may have a color vision deficiency; red-green blindness is the most frequent kind.
Hence avoid relying only on color as the discriminative feature in plots (such as red \vs green lines), but add a second discriminative feature to ease disambiguation.

%------------------------------------------------------------------------
\section{Final copy}

You must include your signed IEEE copyright release form when you submit your finished paper.
We MUST have this form before your paper can be published in the proceedings.

Please direct any questions to the production editor in charge of these proceedings at the IEEE Computer Society Press:
\url{https://www.computer.org/about/contact}.

%%%%%%%%% REFERENCES
{\small
\bibliographystyle{ieee_fullname}
\bibliography{egbib}
}

\end{document}

### 2021-12-12 PaperForReview.tex
updated the contents of the paper

@@
 
-Please follow the steps outlined below when submitting your manuscript to the IEEE Computer Society Press.
-This style guide now has several important modifications (for example, you are no longer warned against the use of sticky tape to attach your artwork to the paper), so all authors should read this new version.
+In today’s age Rising Cardiovascular diseases pegs need of timely diagnosis.
+Apart from regular health checkups an easy & accessible heart health diagnosis solution is needed to prevent mishaps like heart failure.
+One such solution can be - to use the pulse data (similar to ECG data) from wearable health monitoring devices like ‘fitbit’ and analyze that data to detect/diagnose an early warning sign which will enable us to predict the onset of a cardiovascular disease.
+There are various factors which we would have to incorporate such a system – like the power consumption, performance, precision of measurement, desired accuracy of the predictions, etc.
+Now, to build a full ECG abnormality detection and classification mechanism, we would have to choose and implement proper techniques to perform the following actions- data preprocessing, feature extraction, and classification.
+In this project, our main focus would be on the feature extraction step of this design and the usage of wavelet decomposition and analysis techniques - namely - Discrete Wavelet Transform (DWT) and Wavelet Packet Decomposition (WPD) for the feature extraction of ECG signals.
+To see how effective DWT and WPD feature extraction techniques are with respect to different classifiers and what impact do they have on classification accuracy, we have performed the experiment where we vary the scale(5,9,10), mother wavelet type (db4, db6, haar, symlet4), technique( DWT/ WPD) and classifier(SVM, KNN, Fine decision Tree)  to obtain the best working combination of these to maximize the accuracy of the system.
 
 %-------------------------------------------------------------------------
-\subsection{Language}
+\section{Previous Works}
+ECG signals have been traditionally been used to identify cardiovascular abnormalities present in the human body.
+However, pure ECG signals are hard to come by due to the fact that breathing pattern frequencies or electrical interferences corrupt the signal to a great extent.
+As a result, traditional approaches have relied on several pre-processing to actually make the signal worthy of inputting it further into data processing systems.
+Two types of abnormalities corrupt the ECG signals, the first one being the baseline wandering and the next one being
+Baseline wandering is an artifact present in an ECG signal and the causes for baseline wandering are the respiration of the patient and movement while recording the ECG.
+These make the signal hard to interpret as there might be additional frequencies other than the desired heart rate frequency alone.
+The traditional method of removing the baseline wandering is passing the signal into a high pass filter.
+111 paper has a section that talks about many such filtering methods for removing baseline wandering.\\
+Noise in an ECG signal creeps into the ECG recording if the recording pads are not kept properly, Electromyographic noise, and power line interferences.
+These noises can be removed by utilizing a low pass notch filter at 50/60Hz usually.
+But 111 also suggests further developments that have been used to remove the noise more efficiently.\\
+Once the pre-processing is done, different ways are used to extract features which is then fed into the classifier.
+A classic feature extraction method is the Fourier transform method, peak detection( P-QRS-T wave detection).
+Except for the Fourier transform technique, other techniques mainly rely on time-based abnormality detections.
+In the Peak detection technique, the distance between the peaks acts as the main classification basis which is computed as the distance between two R points in the P-QRS-T wave 222.
+A simple method of classification is the Principle component analysis of the signal itself though which the classifier will handle by itself.
+For frequency-based features, performing the FFT followed by the classification based on FFT coefficients is traditionally simple and gives the frequency-based feature classification.
 
-All manuscripts must be in English.
+\section{Proposed Approach}
 
-\subsection{Dual submission}
+In our proposed approach, we use DWT and WPD to perform the feature extraction step instead of using conventional feature extraction methods for ECG signals.
+By using this approach, we aim to optimize the accuracy of the classifier and also provide better (workable) prediction results for low-computational classifiers.
+Apart from this, feature extraction using this method also eliminates the need for noise removal in the data-preprocessing step, thus, we can say that using this technique still keeps the classification process immune to non-linearities of the signal even after removing the noise filtering process in the data-preprocessing step.
+The mentioned Non-linearities of ECG signals like powerline interference will not impact the classification result while using the wavelet transform for feature extraction because when we select the features obtained from wavelet decomposition, we would ignore the data present in trivial frequency bands (containing non-essential data/ noise) and we would not feed that un-essential data to the classifier, thus denoising can be done at the feature extraction step itself.
+Once we are done with the pre-processing, we need to take the DWT/WPD coefficients.
+Compute the classification (Range of signals/energy) as a parameter and give it as an input to the classifiers which were used in the traditional classification methods in order to compare the output results.
+We will perform a comparative study for determining which type of mother wavelet is best suited for this application and up to what scale (levels) of cascading DWT/WPD is useful for this process.
+furthermore, we would also check if the features extracted using wavelet packet decomposition (WPD) give better results than DWT.
+Comparison experiment variables - Mother Wavelet type (Haar, DB, Symlet), Scale of DWT/WPD, Type of Decomposition (WPD or DWT).
 
-Please refer to the author guidelines on the \confName\ \confYear\ web page for a
-discussion of the policy on dual submissions.
-
-\subsection{Paper length}
-Papers, excluding the references section, must be no longer than eight pages in length.
-The references section will not be included in the page count, and there is no limit on the length of the references section.
-For example, a paper of eight pages with two pages of references would have a total length of 10 pages.
+\section{Simulation Results, Comparison & Discussion}
+Papers, excluding the references section, 
 {\bf There will be no extra page charges for \confName\ \confYear.}
@@
 Overlength papers will simply not be reviewed.
-This includes papers where the margins and formatting are deemed to have been significantly altered from those laid down by this style guide.
-Note that this \LaTeX\ guide already sets figure captions and references in a smaller font.
-The reason such papers will not be reviewed is that there is no provision for supervised revisions of manuscripts.
-The reviewing process cannot determine the suitability of the paper for presentation in eight pages if it is reviewed in eleven.
 
 %-------------------------------------------------------------------------
-\subsection{The ruler}
+\section{Conclusion}
 The \LaTeX\ style defines a printed ruler which should be present in the version submitted for review.
@@
 
-Reviewers:
-note that the ruler measurements do not align well with lines in the paper --- this turns out to be very difficult to do well when the paper contains many figures and equations, and, when done, looks ugly.
-Just use fractional references (\eg, this line is $087.5$), although in most cases one would expect that the approximate location will be adequate.
-
-\subsection{Paper ID}
-Make sure that the Paper ID from the submission system is visible in the version submitted for review (replacing the ``*****'' you see in this document).
-If you are using the \LaTeX\ template, \textbf{make sure to update paper ID in the appropriate place in the tex file}.
-
-\subsection{Mathematics}
-
-Please number all of your sections and displayed equations as in these examples:
-\begin{equation}
-  E = m\cdot c^2
-  \label{eq:important}
-\end{equation}
-and
-\begin{equation}
-  v = a\cdot t.
-  \label{eq:also-important}
-\end{equation}
-It is important for readers to be able to refer to any particular equation.
-Just because you did not refer to it in the text does not mean some future reader might not need to refer to it.
-It is cumbersome to have to use circumlocutions like ``the equation second from the top of page 3 column 1''.
-(Note that the ruler will not be present in the final copy, so is not an alternative to equation numbers).
-All authors will benefit from reading Mermin's description of how to write mathematics:
-\url{http://www.pamitc.org/documents/mermin.pdf}.
-
-\subsection{Blind review}
-
-Many authors misunderstand the concept of anonymizing for blind review.
-Blind review does not mean that one must remove citations to one's own work---in fact it is often impossible to review a paper unless the previous citations are known and available.
-
-Blind review means that you do not use the words ``my'' or ``our'' when citing previous work.
-That is all.
-(But see below for tech reports.)
-
-Saying ``this builds on the work of Lucy Smith [1]'' does not say that you are Lucy Smith;
-it says that you are building on her work.
-If you are Smith and Jones, do not say ``as we show in [7]'', say ``as Smith and Jones show in [7]'' and at the end of the paper, include reference 7 as you would any other cited work.
-
-An example of a bad paper just asking to be rejected:
-\begin{quote}
-\begin{center}
-    An analysis of the frobnicatable foo filter.
-\end{center}
-
-   In this paper we present a performance analysis of our previous paper [1], and show it to be inferior to all previously known methods.
-   Why the previous paper was accepted without this analysis is beyond me.
-
-   [1] Removed for blind review
-\end{quote}
-
-An example of an acceptable paper:
-\begin{quote}
-\begin{center}
-     An analysis of the frobnicatable foo filter.
-\end{center}
-
-   In this paper we present a performance analysis of the  paper of Smith \etal [1], and show it to be inferior to all previously known methods.
-   Why the previous paper was accepted without this analysis is beyond me.
-
-   [1] Smith, L and Jones, C. ``The frobnicatable foo filter, a fundamental contribution to human knowledge''. Nature 381(12), 1-213.
-\end{quote}
-
-If you are making a submission to another conference at the same time, which covers similar or overlapping material, you may need to refer to that submission in order to explain the differences, just as you would if you had previously published related work.
-In such cases, include the anonymized parallel submission~\cite{Authors14} as supplemental material and cite it as
-\begin{quote}
-[1] Authors. ``The frobnicatable foo filter'', F\&G 2014 Submission ID 324, Supplied as supplemental material {\tt fg324.pdf}.
-\end{quote}
-
-Finally, you may feel you need to tell the reader that more details can be found elsewhere, and refer them to a technical report.
-For conference submissions, the paper must stand on its own, and not {\em require} the reviewer to go to a tech report for further details.
-Thus, you may say in the body of the paper ``further details may be found in~\cite{Authors14b}''.
-Then submit the tech report as supplemental material.
-Again, you may not assume the reviewers will read this material.
-
-Sometimes your paper is about a problem which you tested using a tool that is widely known to be restricted to a single institution.
-For example, let's say it's 1969, you have solved a key problem on the Apollo lander, and you believe that the CVPR70 audience would like to hear about your
-solution.
-The work is a development of your celebrated 1968 paper entitled ``Zero-g frobnication: How being the only people in the world with access to the Apollo lander source code makes us a wow at parties'', by Zeus \etal.
-
-You can handle this paper like any other.
-Do not write ``We show how to improve our previous work [Anonymous, 1968].
-This time we tested the algorithm on a lunar lander [name of lander removed for blind review]''.
-That would be silly, and would immediately identify the authors.
-Instead write the following:
-\begin{quotation}
-\noindent
-   We describe a system for zero-g frobnication.
-   This system is new because it handles the following cases:
-   A, B.  Previous systems [Zeus et al. 1968] did not  handle case B properly.
-   Ours handles it by including a foo term in the bar integral.
-
-   ...
-
-   The proposed system was integrated with the Apollo lunar lander, and went all the way to the moon, don't you know.
-   It displayed the following behaviours, which show how well we solved cases A and B: ...
-\end{quotation}
-As you can see, the above text follows standard scientific convention, reads better than the first version, and does not explicitly name you as the authors.
-A reviewer might think it likely that the new paper was written by Zeus \etal, but cannot make any decision based on that guess.
-He or she would have to be sure that no other authors could have been contracted to solve problem B.
-\medskip
-
-\noindent
-FAQ\medskip\\
-{\bf Q:} Are acknowledgements OK?\\
-{\bf A:} No.  Leave them for the final copy.\medskip\\
-{\bf Q:} How do I cite my results reported in open challenges?
-{\bf A:} To conform with the double-blind review policy, you can report results of other challenge participants together with your results in your paper.
-For your results, however, you should not identify yourself and should not mention your participation in the challenge.
-Instead present your results referring to the method proposed in your paper and draw conclusions based on the experimental comparison to other results.\medskip\\
-
-\begin{figure}[t]
-  \centering
-  \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
-   %\includegraphics[width=0.8\linewidth]{egfigure.eps}
-
-   \caption{Example of caption.
-   It is set in Roman so that mathematics (always set in Roman: $B \sin A = A \sin B$) may be included without an ugly clash.}
-   \label{fig:onecol}
-\end{figure}
-
-\subsection{Miscellaneous}
-
-\noindent
-Compare the following:\\
-\begin{tabular}{ll}
- \verb'$conf_a$' &  $conf_a$ \\
- \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
-\end{tabular}\\
-See The \TeX book, p165.
-
-The space after \eg, meaning ``for example'', should not be a sentence-ending space.
-So \eg is correct, {\em e.g.} is not.
-The provided \verb'\eg' macro takes care of this.
-
-When citing a multi-author paper, you may save space by using ``et alia'', shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}'' is a complete word).
-If you use the \verb'\etal' macro provided, then you need not worry about double periods when used at the end of a sentence as in Alpher \etal.
-However, use it only when there are three or more authors.
-Thus, the following is correct:
-   ``Frobnication has been trendy lately.
-   It was introduced by Alpher~\cite{Alpher02}, and subsequently developed by
-   Alpher and Fotheringham-Smythe~\cite{Alpher03}, and Alpher \etal~\cite{Alpher04}.''
-
-This is incorrect: ``... subsequently developed by Alpher \etal~\cite{Alpher03} ...'' because reference~\cite{Alpher03} has just two authors.
-
-% Update the cvpr.cls to do the following automatically.
-% For this citation style, keep multiple citations in numerical (not
-% chronological) order, so prefer \cite{Alpher03,Alpher02,Authors14} to
-% \cite{Alpher02,Alpher03,Authors14}.
-
-\begin{figure*}
-  \centering
-  \begin{subfigure}{0.68\linewidth}
-    \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
-    \caption{An example of a subfigure.}
-    \label{fig:short-a}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{0.28\linewidth}
-    \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
-    \caption{Another example of a subfigure.}
-    \label{fig:short-b}
-  \end{subfigure}
-  \caption{Example of a short caption, which should be centered.}
-  \label{fig:short}
-\end{figure*}
-
-%------------------------------------------------------------------------
-\section{Formatting your paper}
-\label{sec:formatting}
-
-All text must be in a two-column format.
-The total allowable size of the text area is $6\frac78$ inches (17.46 cm) wide by $8\frac78$ inches (22.54 cm) high.
-Columns are to be $3\frac14$ inches (8.25 cm) wide, with a $\frac{5}{16}$ inch (0.8 cm) space between them.
-The main title (on the first page) should begin 1 inch (2.54 cm) from the top edge of the page.
-The second and following pages should begin 1 inch (2.54 cm) from the top edge.
-On all pages, the bottom margin should be $1\frac{1}{8}$ inches (2.86 cm) from the bottom edge of the page for $8.5 \times 11$-inch paper;
-for A4 paper, approximately $1\frac{5}{8}$ inches (4.13 cm) from the bottom edge of the
-page.
-
-%-------------------------------------------------------------------------
-\subsection{Margins and page numbering}
-
-All printed material, including text, illustrations, and charts, must be kept
-within a print area $6\frac{7}{8}$ inches (17.46 cm) wide by $8\frac{7}{8}$ inches (22.54 cm)
-high.
-%
-Page numbers should be in the footer, centered and $\frac{3}{4}$ inches from the bottom of the page.
-The review version should have page numbers, yet the final version submitted as camera ready should not show any page numbers.
-The \LaTeX\ template takes care of this when used properly.
-
-%-------------------------------------------------------------------------
-\subsection{Type style and fonts}
-
-Wherever Times is specified, Times Roman may also be used.
-If neither is available on your word processor, please use the font closest in
-appearance to Times to which you have access.
-
-MAIN TITLE.
-Center the title $1\frac{3}{8}$ inches (3.49 cm) from the top edge of the first page.
-The title should be in Times 14-point, boldface type.
-Capitalize the first letter of nouns, pronouns, verbs, adjectives, and adverbs;
-do not capitalize articles, coordinate conjunctions, or prepositions (unless the title begins with such a word).
-Leave two blank lines after the title.
-
-AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title
-and printed in Times 12-point, non-boldface type.
-This information is to be followed by two blank lines.
-
-The ABSTRACT and MAIN TEXT are to be in a two-column format.
-
-MAIN TEXT.
-Type main text in 10-point Times, single-spaced.
-Do NOT use double-spacing.
-All paragraphs should be indented 1 pica (approx.~$\frac{1}{6}$ inch or 0.422 cm).
-Make sure your text is fully justified---that is, flush left and flush right.
-Please do not place any additional blank lines between paragraphs.
-
-Figure and table captions should be 9-point Roman type as in \cref{fig:onecol,fig:short}.
-Short captions should be centred.
-
-\noindent Callouts should be 9-point Helvetica, non-boldface type.
-Initially capitalize only the first word of section titles and first-, second-, and third-order headings.
-
-FIRST-ORDER HEADINGS.
-(For example, {\large \bf 1. Introduction}) should be Times 12-point boldface, initially capitalized, flush left, with one blank line before, and one blank line after.
-
-SECOND-ORDER HEADINGS.
-(For example, { \bf 1.1. Database elements}) should be Times 11-point boldface, initially capitalized, flush left, with one blank line before, and one after.
-If you require a third-order heading (we discourage it), use 10-point Times, boldface, initially capitalized, flush left, preceded by one blank line, followed by a period and your text on the same line.
-
-%-------------------------------------------------------------------------
-\subsection{Footnotes}
-
-Please use footnotes\footnote{This is what a footnote looks like.
-It often distracts the reader from the main flow of the argument.} sparingly.
-Indeed, try to avoid footnotes altogether and include necessary peripheral observations in the text (within parentheses, if you prefer, as in this sentence).
-If you wish to use a footnote, place it at the bottom of the column on the page on which it is referenced.
-Use Times 8-point type, single-spaced.
-
-%-------------------------------------------------------------------------
-\subsection{Cross-references}
-
-For the benefit of author(s) and readers, please use the
-{\small\begin{verbatim}
-  \cref{...}
-\end{verbatim}}  command for cross-referencing to figures, tables, equations, or sections.
-This will automatically insert the appropriate label alongside the cross-reference as in this example:
-\begin{quotation}
-  To see how our method outperforms previous work, please see \cref{fig:onecol} and \cref{tab:example}.
-  It is also possible to refer to multiple targets as once, \eg~to \cref{fig:onecol,fig:short-a}.
-  You may also return to \cref{sec:formatting} or look at \cref{eq:also-important}.
-\end{quotation}
-If you do not wish to abbreviate the label, for example at the beginning of the sentence, you can use the
-{\small\begin{verbatim}
-  \Cref{...}
-\end{verbatim}}
-command. Here is an example:
-\begin{quotation}
-  \Cref{fig:onecol} is also quite important.
-\end{quotation}
-
-%-------------------------------------------------------------------------
-\subsection{References}
-
-List and number all bibliographical references in 9-point Times, single-spaced, at the end of your paper.
-When referenced in the text, enclose the citation number in square brackets, for
-example~\cite{Authors14}.
-Where appropriate, include page numbers and the name(s) of editors of referenced books.
-When you cite multiple papers at once, please make sure that you cite them in numerical order like this \cite{Alpher02,Alpher03,Alpher05,Authors14b,Authors14}.
-If you use the template as advised, this will be taken care of automatically.
-
-\begin{table}
-  \centering
-  \begin{tabular}{@{}lc@{}}
-    \toprule
-    Method & Frobnability \\
-    \midrule
-    Theirs & Frumpy \\
-    Yours & Frobbly \\
-    Ours & Makes one's heart Frob\\
-    \bottomrule
-  \end{tabular}
-  \caption{Results.   Ours is better.}
-  \label{tab:example}
-\end{table}
-
-%-------------------------------------------------------------------------
-\subsection{Illustrations, graphs, and photographs}
-
-All graphics should be centered.
-In \LaTeX, avoid using the \texttt{center} environment for this purpose, as this adds potentially unwanted whitespace.
-Instead use
-{\small\begin{verbatim}
-  \centering
-\end{verbatim}}
-at the beginning of your figure.
-Please ensure that any point you wish to make is resolvable in a printed copy of the paper.
-Resize fonts in figures to match the font in the body text, and choose line widths that render effectively in print.
-Readers (and reviewers), even of an electronic copy, may choose to print your paper in order to read it.
-You cannot insist that they do otherwise, and therefore must not assume that they can zoom in to see tiny details on a graphic.
-
-When placing figures in \LaTeX, it's almost always best to use \verb+\includegraphics+, and to specify the figure width as a multiple of the line width as in the example below
-{\small\begin{verbatim}
-   \usepackage{graphicx} ...
-   \includegraphics[width=0.8\linewidth]
-                   {myfile.pdf}
-\end{verbatim}
-}
-
-%-------------------------------------------------------------------------
-\subsection{Color}
-
-Please refer to the author guidelines on the \confName\ \confYear\ web page for a discussion of the use of color in your document.
-
-If you use color in your plots, please keep in mind that a significant subset of reviewers and readers may have a color vision deficiency; red-green blindness is the most frequent kind.
-Hence avoid relying only on color as the discriminative feature in plots (such as red \vs green lines), but add a second discriminative feature to ease disambiguation.
-
-%------------------------------------------------------------------------
 \section{Final copy}

### 2021-12-12 PaperForReview.tex
added images and updates

@@
 \usepackage{booktabs}
+\usepackage{multirow}
 
@@
 %%%%%%%%% PAPER ID  - PLEASE UPDATE
-\def\cvprPaperID{*****} % *** Enter the CVPR Paper ID here
+\def\cvprPaperID{ Group 7} % *** Enter the CVPR Paper ID here
 \def\confName{ECE 251C}
@@
 UCSD\\
-Institution1 address\\
 {\tt\small hsankar@ucsd.edu}
-% For a paper whose authors are all at the same institution,
-% omit the following lines up until the closing ``}''.
-% Additional authors and addresses can be added with ``\and'',
-% just like the second author.
-% To save space, use either the email address or home page, not both
 \and
@@
 UCSD\\
-First line of institution2 address\\
 {\tt\small akhare@ucsd.edu}
@@
 \label{sec:intro}
-
 In today’s age Rising Cardiovascular diseases pegs need of timely diagnosis.
-Apart from regular health checkups an easy & accessible heart health diagnosis solution is needed to prevent mishaps like heart failure.
-One such solution can be - to use the pulse data (similar to ECG data) from wearable health monitoring devices like ‘fitbit’ and analyze that data to detect/diagnose an early warning sign which will enable us to predict the onset of a cardiovascular disease.
+Apart from regular health checkups an easy \& accessible heart health diagnosis solution is needed to prevent mishaps like heart failure.One such solution can be - to use the pulse data (similar to ECG data) from wearable health monitoring devices like fitbit and analyze that data to detect/diagnose an early warning sign which will enable us to predict the onset of a cardiovascular disease.
 There are various factors which we would have to incorporate such a system – like the power consumption, performance, precision of measurement, desired accuracy of the predictions, etc.
@@
 As a result, traditional approaches have relied on several pre-processing to actually make the signal worthy of inputting it further into data processing systems.
-Two types of abnormalities corrupt the ECG signals, the first one being the baseline wandering and the next one being
+Two types of abnormalities corrupt the ECG signals, the first one being the baseline wandering and the next one being noise.
+
+\begin{figure}[htbp]
+\includegraphics[width=8cm]{traditional.png}
+\end{figure}
+
 Baseline wandering is an artifact present in an ECG signal and the causes for baseline wandering are the respiration of the patient and movement while recording the ECG.
@@
 The traditional method of removing the baseline wandering is passing the signal into a high pass filter.
-111 paper has a section that talks about many such filtering methods for removing baseline wandering.\\
+111 paper has a section that talks about many such filtering methods for removing baseline wandering.
 Noise in an ECG signal creeps into the ECG recording if the recording pads are not kept properly, Electromyographic noise, and power line interferences.
 These noises can be removed by utilizing a low pass notch filter at 50/60Hz usually.
-But 111 also suggests further developments that have been used to remove the noise more efficiently.\\
+But [Cite 1] also suggests further developments that have been used to remove the noise more efficiently.
 Once the pre-processing is done, different ways are used to extract features which is then fed into the classifier.
@@
 Except for the Fourier transform technique, other techniques mainly rely on time-based abnormality detections.
-In the Peak detection technique, the distance between the peaks acts as the main classification basis which is computed as the distance between two R points in the P-QRS-T wave 222.
+In the Peak detection technique, the distance between the peaks acts as the main classification basis which is computed as the distance between two R points in the P-QRS-T wave [cite 2].
 A simple method of classification is the Principle component analysis of the signal itself though which the classifier will handle by itself.
@@
 
-In our proposed approach, we use DWT and WPD to perform the feature extraction step instead of using conventional feature extraction methods for ECG signals.
-By using this approach, we aim to optimize the accuracy of the classifier and also provide better (workable) prediction results for low-computational classifiers.
-Apart from this, feature extraction using this method also eliminates the need for noise removal in the data-preprocessing step, thus, we can say that using this technique still keeps the classification process immune to non-linearities of the signal even after removing the noise filtering process in the data-preprocessing step.
-The mentioned Non-linearities of ECG signals like powerline interference will not impact the classification result while using the wavelet transform for feature extraction because when we select the features obtained from wavelet decomposition, we would ignore the data present in trivial frequency bands (containing non-essential data/ noise) and we would not feed that un-essential data to the classifier, thus denoising can be done at the feature extraction step itself.
-Once we are done with the pre-processing, we need to take the DWT/WPD coefficients.
-Compute the classification (Range of signals/energy) as a parameter and give it as an input to the classifiers which were used in the traditional classification methods in order to compare the output results.
+In our proposed approach, we use DWT and WPD to perform the feature extraction step instead of using conventional feature extraction methods for ECG signals.\\
+The concept of wavelet transform can be explained simply as decomposing or transforming a signal into multiple orthogonal basis functions.
+However, the difference between Fourier transform and wavelet transform or decomposition is the fact that Fourier provides resolution in frequency domain alone whereas wavelets act like a hierarchical setup of basis functions across time, where we can point out which frequency is present in what time to a considerable extent.
+As shown in the figure, DWT decomposes the signal into high pass and low pass sections and as the scales go beyond 1, it takes the low pass functions alone to further look deep into its components.
+The main difference between discrete Wavelet transform and Wavelet packet decomposition stems from what these techniques decompose and look further.
+Wavelet packet decomposition takes both the lowpass and highpass components and decomposes them into further lowpass and highpass, thus providing a higher overall resolution of the complete signal.
+Thus, Wavelet packet transform results in equal-width sub-band filtering of signals as opposed to the coarser band filtering found in DWT.
+\begin{figure}[tbp]
+\includegraphics[width=8cm]{dwt.JPG}
+\end{figure}
+\begin{figure}[tbp]
+\includegraphics[width=8cm]{wpd.JPG}
+\end{figure}
+\cite{Alpher01,Alpher02,Alpher03,Alpher04, Alpher05}
+Comparison experiment variables - Mother Wavelet type (Haar, DB, Symlet), Scale of DWT/WPD, Type of Decomposition (WPD or DWT).\\
+By using these two approaches, we aim to optimize the accuracy of the classifier and also provide better (workable) prediction results for low-computational classifiers.
+Apart from this, feature extraction using this method also eliminates the need for noise removal in the data pre-processing step, , thus, we can say that using this technique still keeps the classification process immune to non-linearities of the signal even after removing the noise filtering process in the data pre-processing step.
+\begin{figure}[htbp]
+\includegraphics[width=8cm]{proposed.png}
+\end{figure}
+
+The mentioned Non-linearities of ECG signals like powerline interference will not impact the classification result while using the wavelet transform for feature extraction because when we select the features obtained from wavelet decomposition, we would ignore the data present in trivial frequency bands (containing non-essential data/ noise) and we would not feed that non-essential data to the classifier, thus denoising can be done at the feature extraction step itself.
+Once we are done with the basic data pre-processing like baseline wandering removal, we need to compute the DWT/WPD coefficients for all scales.
+After calculating the DWT approximate and detailed coefficients, we use the selected scales of data that are best suited for our classification requirements.
+We take the coefficients of the selected scale and we divide the coefficients’ data into smaller chunks and calculate the energy content of each such chunk.
+Once we have the energy values of these chunks, we calculate the distribution of energy in these chunks and give that as an input parameter to the classifier.
+The classifier then takes these parameters as input from every such scale which was selected by us and runs the respective classification algorithm.
+We then compare the obtained results from the ones obtained from the traditional classification methods.
+In WPD, for feature extraction purposes, 333 had a good idea about using all the coefficients of the last scale of decomposition.
+Using this idea as the base, we checked the energy percentage distribution across coefficients and found that most of the energy was concentrated around 1-16 indexes.
+We have used these distribution parameters as our feature extraction pivots and have classified the healthy and unhealthy data sets.
 We will perform a comparative study for determining which type of mother wavelet is best suited for this application and up to what scale (levels) of cascading DWT/WPD is useful for this process.
 furthermore, we would also check if the features extracted using wavelet packet decomposition (WPD) give better results than DWT.
-Comparison experiment variables - Mother Wavelet type (Haar, DB, Symlet), Scale of DWT/WPD, Type of Decomposition (WPD or DWT).
 
-\section{Simulation Results, Comparison & Discussion}
-Papers, excluding the references section, 
-{\bf There will be no extra page charges for \confName\ \confYear.}
+\section{Simulation Results, Comparison \& Discussion}
+\subsection{Simulation Environment}
+The simulation environment consisted of us extracting the data from [repo link] which consist of patients with various heart arrhythmia.
+Each data was of length 30 mins.
+So in order to reduce the computationally convenient and add more data, we chose to take individual sections of the ECG that had problems as our unhealthy signal and portions of ECG signal that did not have any arrhythmia as healthy ones.
+Each of the unhealthy and healthy signal length lasted for a minute.
+We then used the input signal after doing pre-processing to get the features that we have mentioned above and fed it into the classifiers to see the performance.
 
-Overlength papers will simply not be reviewed.
+\subsection{Simulation Results} 
+We ran the simulation for 5 types of wavelets and depending on the type of method used DWT/WPD, we chose scales.
+Scales of 9 and 10 were used for DWT and scales of 5 and 9 were used for WPD.
+The tables X and Table Y shows us the accuracy results of the classification for a particular type of classifier.
+
+\begin{table}[htbp]
+\scalebox{0.9}{
+\begin{tabular}{|c|cc|cc|}
+\hline
+\begin{tabular}[c]{@{}c@{}}KNN \\ Classifier\end{tabular} & \multicolumn{2}{c|}{Scale : 9}                                                            & \multicolumn{2}{c|}{Scale : 10}                                                           \\ \hline
+Method                                                    & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\end{tabular}} & Accuracy & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\end{tabular}} & Accuracy \\ \hline
+\multirow{5}{*}{DWT}                                      & \multicolumn{1}{c|}{Db2}                                                       & 80       & \multicolumn{1}{c|}{Db2}                                                       & 86       \\ \cline{2-5}
+                                                          & \multicolumn{1}{c|}{Db4}                                                       & 72       & \multicolumn{1}{c|}{Db4}                                                       & 79       \\ \cline{2-5}
+                                                          & \multicolumn{1}{c|}{Db6}                                                       & 75       & \multicolumn{1}{c|}{Db6}                                                       & 84       \\ \cline{2-5}
+                                                          & \multicolumn{1}{c|}{Sym4}                                                      & 80       & \multicolumn{1}{c|}{Sym4}                                                      & 81       \\ \cline{2-5}
+                                                          & \multicolumn{1}{c|}{Haar}                                                      & 72       & \multicolumn{1}{c|}{Haar}                                                      & 70       \\ \hline
+\end{tabular}}
+\end{table}
+
+%%----------------------------%%
+
+\begin{table}[htbp]
+\scalebox{0.9}{
+\begin{tabular}{|c|cc|cc|}
+\hline
+\begin{tabular}[c]{@{}c@{}}KNN \\ Classifier\end{tabular} & \multicolumn{2}{c|}{Scale : 5}                                                            & \multicolumn{2}{c|}{Scale : 9}                                                            \\ \hline
+Method                                                    & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\end{tabular}} & Accuracy & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\end{tabular}} & Accuracy \\ \hline
+\multirow{5}{*}{WPD}                                      & \multicolumn{1}{c|}{Db2}                                                       & 80       & \multicolumn{1}{c|}{Db2}                                                       & 86       \\ \cline{2-5}
+                                                          & \multicolumn{1}{c|}{Db4}                                                       & 78       & \multicolumn{1}{c|}{Db4}                                                       & 84       \\ \cline{2-5}
+                                                          & \multicolumn{1}{c|}{Db6}                                                       & 84       & \multicolumn{1}{c|}{Db6}                                                       & 88       \\ \cline{2-5}
+                                                          & \multicolumn{1}{c|}{Sym4}                                                      & 80       & \multicolumn{1}{c|}{Sym4}                                                      & 82       \\ \cline{2-5}
+                                                          & \multicolumn{1}{c|}{Haar}                                                      & 78       & \multicolumn{1}{c|}{Haar}                                                      & 78       \\ \hline
+\end{tabular}}
+\end{table}
+
+%%----------------------------%%
+
+\begin{table}[htbp]
+\scalebox{0.7}{
+\begin{tabular}{|c|c|c|cl|c|c|}
+\hline
+SVM Classifier                                                   & scale 9  & scale10  & \multicolumn{2}{c|}{}                                                                & scale 5  & scale9   \\ \hline
+DWT                                                              & Accuracy & Accuracy & \multicolumn{2}{c|}{WPD}                                                             & Accuracy & Accuracy \\ \hline
+\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet \\ Db6\end{tabular} & 81.4     & 85.3     & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\\ Db6\end{tabular}} & 83       & 87       \\ \hline
+\end{tabular}}
+\end{table}
+
+%%----------------------------%%
+
+\begin{table}[htbp]
+\scalebox{0.7}{
+\begin{tabular}{|l|l|l|ll|l|l|}
+\hline
+\begin{tabular}[c]{@{}l@{}}Decision Tree \\ Classifier\end{tabular} & scale 9  & scale10  & \multicolumn{2}{l|}{}                                                                 & scale 5  & scale9   \\ \hline
+DWT                                                                 & Accuracy & Accuracy & \multicolumn{2}{l|}{WPD}                                                              & Accuracy & Accuracy \\ \hline
+\begin{tabular}[c]{@{}l@{}}Mother \\ Wavelet \\ Db6\end{tabular}    & 72       & 84       & \multicolumn{2}{l|}{\begin{tabular}[c]{@{}l@{}}Mother \\ Wavelet \\ Db6\end{tabular}} & 82       & 83       \\ \hline
+\end{tabular}}
+\end{table}
+
+As we can see from the results, in general of performance of WPD is better than DWT which is expected because of the fact that WPD will have more information when compared to the DWT of the same scale.
 
@@
 \section{Conclusion}
-The \LaTeX\ style defines a printed ruler which should be present in the version submitted for review.
-The ruler is provided in order that reviewers may comment on particular lines in the paper without circumlocution.
-If you are preparing a document using a non-\LaTeX\ document preparation system, please arrange for an equivalent ruler to appear on the final output pages.
-The presence or absence of the ruler should not change the appearance of any other content on the page.
-The camera-ready copy should not contain a ruler.
-(\LaTeX\ users may use options of cvpr.sty to switch between different versions.)
-
-\section{Final copy}
-
-You must include your signed IEEE copyright release form when you submit your finished paper.
-We MUST have this form before your paper can be published in the proceedings.
-
-Please direct any questions to the production editor in charge of these proceedings at the IEEE Computer Society Press:
-\url{https://www.computer.org/about/contact}.
+In conclusion, we can deduce that the proposed method of feature extraction using DWT and DWD is much more computationally efficient ( large scale data) than conventional methods (peak detection, QRS Detection), and by using this method we are able to reduce the number of features/dimensions of the data being fed to the classifier, thus it drastically reduces the latency of the classifier.
+Using wavelet decomposition also doesn’t negatively affect the amount of useful information sent to the classifier that is used for classification.
+We were also able to improve the accuracy of the classification results to a good extent. We con
+The filtering step in the pre-processing stage is being done at the feature extraction step itself instead of using separate stages.
+This resulted in the system being more time-efficient and the complexity of the overall system was hence reduced.
+Further, we can say that this method can be easily applied in multi-class classification for detecting different types of abnormalities like arrhythmia, atrial flutter, fibrillation, etc to make the diagnosis more precise.
 
@@
 
+\section*{Miscellaneous Links}
+\begin{itemize}
+    \item \href{https://physionet.org/content/mitdb/1.0.0/}{Open Source Database}
+    \item \href{https://github.com/abhishekkhare1998/EcgClassificationProject}{Repository}
+\end{itemize}
+
 \end{document}

### 2021-12-12 PaperForReview.tex
added more data and images

@@
 \begin{abstract}
-   The ABSTRACT is to be in fully justified italicized text, at the top of the left-hand column, below the author and affiliation information.
-   Use the word ``Abstract'' as the title, in 12-point Times, boldface type, centered relative to the column, initially capitalized.
-   The abstract is to be in 10-point, single-spaced type.
-   Leave two blank lines after the Abstract, then begin the main text.
-   Look at previous CVPR abstracts to get a feel for style and length.
+   Electrocardiogram (ECG) are signals which captures the mechanical activity of heart.
+These signals are usually fairly periodic and they provide the first-hand insights into any cardiovascular abnormalities or diseases that a person might have in the future.
+We can take advantage of the periodicity of these signals to classify various heart related conditions depending upon the variations which these ECG signals undergo.This project discusses two such methods to perform binary classification of ECG signals between healthy and unhealthy ones.
+Specifically, we take the concepts of wavelets and wavelet packet decomposition to extract features from these signals by performing mathematical operations on them and feed that extracted feature data to the Machine Learning based classifier.
+Finally, we would want to come to a conclusion about the most suitable classifier, decomposition level and wavelet combination.\\
+   \providecommand{\keywords}[1]
+{
+  \small	
+  \textbf{\textit{Keywords---}} #1
+}
+\keywords{DWT - Discrete Wavelet Transform, WPD - Wavelet Packet Decomposition, ECG - Electrocardiogram, SVM - Support Vector Machine, KNN - k-Nearest Neighbours, Db - Daubechies}
 \end{abstract}
@@
 To see how effective DWT and WPD feature extraction techniques are with respect to different classifiers and what impact do they have on classification accuracy, we have performed the experiment where we vary the scale(5,9,10), mother wavelet type (db4, db6, haar, symlet4), technique( DWT/ WPD) and classifier(SVM, KNN, Fine decision Tree)  to obtain the best working combination of these to maximize the accuracy of the system.
-
+Figure 1 shows the difference between a Healthy ECG signal and an unhealthy ECG signal that has arrhythmia.
+\begin{figure}[h!]
+\includegraphics[width=8cm]{healthyVSunhealthy.JPG}
+\caption{Healthy vs Unhealthy  ECG Signal Example}
+\label{Comparison}
+\end{figure}
 %-------------------------------------------------------------------------
@@
 \includegraphics[width=8cm]{traditional.png}
+\caption{Conventional method process}
+\label{Conventional Method}
 \end{figure}
@@
 The traditional method of removing the baseline wandering is passing the signal into a high pass filter.
-111 paper has a section that talks about many such filtering methods for removing baseline wandering.
+111 paper has a section that talks about many such filtering methods for removing baseline wandering.\\
+\begin{figure}[htbp]
+\includegraphics[width=8cm]{ConventionalResult.jpg}
+\caption{Conventional method's accuracy}
+\label{Classification using conventional approach}
+\end{figure}
 Noise in an ECG signal creeps into the ECG recording if the recording pads are not kept properly, Electromyographic noise, and power line interferences.
@@
 A simple method of classification is the Principle component analysis of the signal itself though which the classifier will handle by itself.
-For frequency-based features, performing the FFT followed by the classification based on FFT coefficients is traditionally simple and gives the frequency-based feature classification.
+For frequency-based features, performing the FFT followed by the classification based on FFT coefficients is traditionally simple and gives the frequency-based feature classification.\\
+We simulated the simplistic conventional model and we have presented the classifier output accuracy as shown in Figure 3.
 
@@
 However, the difference between Fourier transform and wavelet transform or decomposition is the fact that Fourier provides resolution in frequency domain alone whereas wavelets act like a hierarchical setup of basis functions across time, where we can point out which frequency is present in what time to a considerable extent.
-As shown in the figure, DWT decomposes the signal into high pass and low pass sections and as the scales go beyond 1, it takes the low pass functions alone to further look deep into its components.
-The main difference between discrete Wavelet transform and Wavelet packet decomposition stems from what these techniques decompose and look further.
+As shown in the figure 4, DWT decomposes the signal into high pass and low pass sections and as the scales go beyond 1, it takes the low pass functions alone to further look deep into its components.
+As seen in Figure 5, the main difference between discrete Wavelet transform and Wavelet packet decomposition stems from what these techniques decompose and look further.
 Wavelet packet decomposition takes both the lowpass and highpass components and decomposes them into further lowpass and highpass, thus providing a higher overall resolution of the complete signal.
 Thus, Wavelet packet transform results in equal-width sub-band filtering of signals as opposed to the coarser band filtering found in DWT.
-\begin{figure}[tbp]
+
+\begin{figure}[htbp]
 \includegraphics[width=8cm]{dwt.JPG}
+\caption{frequency domain scaling representation of DWT}
+\label{DWT}
 \end{figure}
-\begin{figure}[tbp]
+
+\begin{figure}[htbp]
 \includegraphics[width=8cm]{wpd.JPG}
+\caption{frequency domain scaling representation of WPD}
+\label{WPD}
 \end{figure}
+
 \cite{Alpher01,Alpher02,Alpher03,Alpher04, Alpher05}
@@
 \includegraphics[width=8cm]{proposed.png}
+\caption{Proposed Algorithm Flow}
+\label{Proposed Algorithm}
 \end{figure}
@@
 \section{Simulation Results, Comparison \& Discussion}
+In this section, we will discuss the experimental setup , results and discuss the observations from the classifier output.
 \subsection{Simulation Environment}
@@
 Scales of 9 and 10 were used for DWT and scales of 5 and 9 were used for WPD.
-The tables X and Table Y shows us the accuracy results of the classification for a particular type of classifier.
+The Table 1 and Table 2 shows us the accuracy results of the classification for a particular type of classifier.
 
+\begin{figure}[h!]
+\includegraphics[width=8.5cm, height=6cm]{DWT_example.png}
+\caption{Decomposition of original signal using DWT}
+\label{DWT Simulation}
+\end{figure}
+As we can see in figure 7, the arrhythmia in the ECG signal can easily be detected using scales 6,7, and 8 of DWT decomposition, and the rest of the coefficients in other scales are not of much use to our classifier, so we would use the data only from the given scales for classification in case of DWT.\\
+\begin{figure}[h!]
+\includegraphics[width=8.5cm, height=6cm]{WPD_example.png}
+\caption{Decomposition of original signal using WPD \\ node - (7,1)}
+\label{WPD Simulation}
+\end{figure}
+Similarly in Figure 8, we visualize wavelet packet decomposition and observe that a few nodes (here - 7,1) show much more pronounced differentiation between normal and abnormal ECG signals, and similar to our method in DWT feature extraction, here we would only consider a few important nodes which are supposed to give better results.\\
+\begin{figure}[h!]
+\includegraphics[width=8.5cm]{DWT_result_db2.jpg}
+\caption{Classifier Output when using DWT \\ (here mother wavelet Db2) based feature extraction}
+\label{DWT Simulation}
+\end{figure}
+
+\begin{figure}[htbp]
+\includegraphics[width=8.5cm]{WPD_result_db6.jpg}
+\caption{Classifier Output when using WPD \\ (here mother wavelet Db6) based feature extraction}
+\label{WPD Simulation}
+\end{figure}
+Using these two techniques we give this extracted features data as an input to various classifiers which produce the confusion matrices as shown in figures 9 \& 10.
+From this result, we see that we get around 85\% \& 88\% classification accuracy.
+Similarly, we ran this simulation setup for various mother wavelets with varying scales on different classifiers and compared the results to select the best possible combination - (feature extraction type, wavelet, scale, classifier)
+\subsection{Comparison \& Discussion}
+
+As we can see from the results, in general the classification accuracy of using WPD feature extraction is better than using DWT even at lower scales.
+This result was expected because of the fact that WPD will have more information when compared to the DWT of the same scale.
+\\
+As observed in Table 1 , when using DWT as feature extraction method we have to use scale up to 9 \& 10 for getting a reasonable classification result.
+When we chose mother wavelets like Haar and Symlets, we see the classification accuracy is not as good as Daubechies mother wavelets.
+Specifically db2 provides the best accuracy at both scales 9 and 10, with accuracy reaching 86\% at scale 10.
+Now, observing the accuracy results improvement across scales, we see that there is a decent increment in accuracy when we using higher scales like scales 8-10, but if we further increase the number of levels, the classification accuracy would start saturating due to the non-availability of meaningful samples as it is too down-sampled.
+
+%% --------------%%
 \begin{table}[htbp]
@@
 \end{tabular}}
+\caption{Accuracy Comparison for DWT across varying \\ Mother Wavelets and Scales using KNN classifier}
+\label{tab:caption}
 \end{table}
+ In Table 2, we see classification accuracy results while using WPD as the feature extraction method.
+Here, for scales comparison we are using scale 5 and scale 9 because we found that even scale 5 was able to match the classification accuracy of DWT.
+We tried to analyse the accuracy increment across higher scales.
+So we decided to compare scale 9 which was used in DWT to have a credible benchmark for our comparison.
+As we can interpret from this table the Haar and Symlets mother wavelets results are similar irrespective of the feature extraction method, however we see that the performance of db6 yields the best results when using WPD and almost matches with db2's performance in DWT even when using just scale 5 in WPD.\\
 
@@
 \end{tabular}}
+\caption{Accuracy Comparison for WPD across varying \\ Mother Wavelets and Scales using KNN classifier}
+\label{tab:caption}
 \end{table}
-
+%% --------------------------%%
+ Overall, we can say that db6 mother wavelet performs the best for our application closely followed by db2.
+Also, for our particular application, when using DWT feature extraction method, we have to use higher scales like scales 9-10 to achieve decent accuracy.
+While we would only need minimum decomposition scale 5 to achieve same accuracy in WPD method.
+However, we can further increase the accuracy of WPD method by increasing the decomposition levels up to 9-10.
 %%----------------------------%%
@@
 \end{tabular}}
+\caption{SVM - Classification Accuracy : DWT vs WPD}
+\label{tab:caption}
 \end{table}
@@
 \end{tabular}}
+\caption{Decision Tree - Classification Accuracy :- \\ DWT vs WPD}
+\label{tab:caption}
 \end{table}
-
-As we can see from the results, in general of performance of WPD is better than DWT which is expected because of the fact that WPD will have more information when compared to the DWT of the same scale.
+%%------------------------------%%
+We also need to check these methods efficacy with respect to different classifiers like decision tree and Support Vector Machine (SVM), K- Nearest Neighbours (KNN).
+So we decided to take the best mother wavelet (in our case db6) and optimal decomposition scales of 5 \& 9 for WPD and 9 \& 10 for DWT respectively.
+since we had already used KNN as the base classifier for table 1 \& 2 , hence we decided to compare the accuracy results using SVM and Decision Tree.\\
+As we can see from Tables 3 \& 4, the decision tree classifiers performs the best of for scale 10 DWT feature extraction.
+whereas, with regards to scale 5, the WPD's accuracy results are equivalent to DWT's best results.
+For SVM Classifiers, WPDs performance across scales is superior to DWT's performance.\\
+As a result, we can say that when we are using KNN or SVM, WPD feature extraction is superior and is decomposition level agnostic, whereas in decision tree's case for higher levels of decomposition one can opt for DWT method of decomposition.\\
+As we can infer from the above discussion, using wavelets can simplify our task of classification of ECG signals to a great extent.
+However, the method also comes with an additional cost of Wavelet Packet Transform time overhead.
+Also, for non-binary classification purposes we need purely adaptive method of scale and mother wavelet selection which is out of the scope of this project.
 
@@
 \section{Conclusion}
-In conclusion, we can deduce that the proposed method of feature extraction using DWT and DWD is much more computationally efficient ( large scale data) than conventional methods (peak detection, QRS Detection), and by using this method we are able to reduce the number of features/dimensions of the data being fed to the classifier, thus it drastically reduces the latency of the classifier.
+In conclusion, we can deduce that the proposed method of feature extraction using DWT and DWD is much more computationally efficient ( for large scale data) than conventional methods (peak detection, QRS Detection), and by using this method we are able to reduce the number of features/dimensions of the data being fed to the classifier, thus it drastically reduces the latency of the classifier.
 Using wavelet decomposition also doesn’t negatively affect the amount of useful information sent to the classifier that is used for classification.
@@
 \begin{itemize}
-    \item \href{https://physionet.org/content/mitdb/1.0.0/}{Open Source Database}
-    \item \href{https://github.com/abhishekkhare1998/EcgClassificationProject}{Repository}
+    \item \href{https://physionet.org/content/mitdb/1.0.0/}{Open Source MIT-BIH Database}
+    \item \href{https://github.com/abhishekkhare1998/EcgClassificationProject}{Project Repository}
+    \item \href{https://github.com/abhishekkhare1998/EcgClassWaveletPaper}{Report Repository}
 \end{itemize}

### 2021-12-12 PaperForReview.tex
Final Updates

@@
 %%%%%%%%% PAPER ID  - PLEASE UPDATE
-\def\cvprPaperID{ Group 7} % *** Enter the CVPR Paper ID here
+\def\cvprPaperID{Group 7} % *** Enter the CVPR Paper ID here
 \def\confName{ECE 251C}
@@
 %%%%%%%%% TITLE - PLEASE UPDATE
-\title{Feature Extraction of ECG signals using Wavelet transform \confName~Final Project report}
+\title{Feature Extraction of ECG signals using Wavelet transform}
 
-\author{Hari Prasad Sankar\\
-UCSD\\
-{\tt\small hsankar@ucsd.edu}
-\and
-Abhishek Khare\\
-UCSD\\
-{\tt\small akhare@ucsd.edu}
-}
 \maketitle
@@
 \begin{abstract}
-   Electrocardiogram (ECG) are signals which captures the mechanical activity of heart.
-These signals are usually fairly periodic and they provide the first-hand insights into any cardiovascular abnormalities or diseases that a person might have in the future.
-We can take advantage of the periodicity of these signals to classify various heart related conditions depending upon the variations which these ECG signals undergo.This project discusses two such methods to perform binary classification of ECG signals between healthy and unhealthy ones.
-Specifically, we take the concepts of wavelets and wavelet packet decomposition to extract features from these signals by performing mathematical operations on them and feed that extracted feature data to the Machine Learning based classifier.
-Finally, we would want to come to a conclusion about the most suitable classifier, decomposition level and wavelet combination.\\
+Electrocardiograms (ECG) are signals which capture the mechanical activity of the heart.
+These signals are usually fairly periodic and they provide first-hand insights into any cardiovascular abnormalities or diseases that a person might have in the future.
+We can take advantage of the periodicity of these signals to classify various heart-related conditions depending upon the variations which these ECG signals undergo.
+This project discusses two such methods to perform binary classification of ECG signals between healthy and unhealthy ones.
+Specifically, we take the concepts of wavelets and wavelet packet decomposition to extract features from these signals by performing mathematical operations on them and feeding that extracted feature data to the Machine Learning based classifier.
+Finally, we would want to come to a conclusion about the most suitable classifier, decomposition level, and wavelet combination.\\
    \providecommand{\keywords}[1]
@@
 \label{sec:intro}
-In today’s age Rising Cardiovascular diseases pegs need of timely diagnosis.
-Apart from regular health checkups an easy \& accessible heart health diagnosis solution is needed to prevent mishaps like heart failure.One such solution can be - to use the pulse data (similar to ECG data) from wearable health monitoring devices like fitbit and analyze that data to detect/diagnose an early warning sign which will enable us to predict the onset of a cardiovascular disease.
+In today’s age Rising Cardiovascular diseases, pegs need of timely diagnosis.
+Apart from regular health checkups, an easy \& accessible heart health diagnosis solution is needed to prevent mishaps like heart failure.
+One such solution can be - to use the pulse data (similar to ECG data) from wearable health monitoring devices like Fitbit and analyze that data to detect/diagnose an early warning sign which will enable us to predict the onset of cardiovascular disease.
 There are various factors which we would have to incorporate such a system – like the power consumption, performance, precision of measurement, desired accuracy of the predictions, etc.
@@
 These make the signal hard to interpret as there might be additional frequencies other than the desired heart rate frequency alone.
-The traditional method of removing the baseline wandering is passing the signal into a high pass filter.
-111 paper has a section that talks about many such filtering methods for removing baseline wandering.\\
+The traditional method of removing the baseline wandering is passing the signal into a high pass filter\cite{Alpher05}.
+Paper has a section that talks about many such filtering methods for removing baseline wandering.\\
 \begin{figure}[htbp]
@@
  But [Cite 1] also suggests further developments that have been used to remove the noise more efficiently.
-Once the pre-processing is done, different ways are used to extract features which is then fed into the classifier.
+Once the pre-processing is done, different ways are used to extract features which are then fed into the classifier.
 A classic feature extraction method is the Fourier transform method, peak detection( P-QRS-T wave detection).
@@
 In our proposed approach, we use DWT and WPD to perform the feature extraction step instead of using conventional feature extraction methods for ECG signals.\\
-The concept of wavelet transform can be explained simply as decomposing or transforming a signal into multiple orthogonal basis functions.
+The concept of wavelet transform can be explained simply as decomposing or transforming a signal into multiple orthogonal basis functions\cite{Alpher04}.
 However, the difference between Fourier transform and wavelet transform or decomposition is the fact that Fourier provides resolution in frequency domain alone whereas wavelets act like a hierarchical setup of basis functions across time, where we can point out which frequency is present in what time to a considerable extent.
-As shown in the figure 4, DWT decomposes the signal into high pass and low pass sections and as the scales go beyond 1, it takes the low pass functions alone to further look deep into its components.
+As shown in figure 4, DWT decomposes the signal into high pass and low pass sections and as the scales go beyond 1, it takes the low pass functions alone to further look deep into its components.
 As seen in Figure 5, the main difference between discrete Wavelet transform and Wavelet packet decomposition stems from what these techniques decompose and look further.
@@
 Thus, Wavelet packet transform results in equal-width sub-band filtering of signals as opposed to the coarser band filtering found in DWT.
-
 \begin{figure}[htbp]
@@
 
-\cite{Alpher01,Alpher02,Alpher03,Alpher04, Alpher05}
-Comparison experiment variables - Mother Wavelet type (Haar, DB, Symlet), Scale of DWT/WPD, Type of Decomposition (WPD or DWT).\\
-By using these two approaches, we aim to optimize the accuracy of the classifier and also provide better (workable) prediction results for low-computational classifiers.
-Apart from this, feature extraction using this method also eliminates the need for noise removal in the data pre-processing step, , thus, we can say that using this technique still keeps the classification process immune to non-linearities of the signal even after removing the noise filtering process in the data pre-processing step.
+%\cite{Alpher01,Alpher02,Alpher03,Alpher04, Alpher05}
+%Comparison experiment variables - Mother Wavelet type (Haar, DB, Symlet), Scale of DWT/WPD, Type of Decomposition (WPD or DWT).\\
+By using these two approaches, we aim to optimize the accuracy of the classifier and also provide better (workable) prediction results for low-computational classifiers \cite{Alpher02}.
+Paper \cite{Alpher03}  takes DWT and uses a similar classification approach.
+Apart from the optimization that this method provides, feature extraction using this method also eliminates the need for noise removal in the data pre-processing step, thus, we can say that using this technique still keeps the classification process immune to non-linearities of the signal even after removing the noise filtering process in the data pre-processing step.
 \begin{figure}[htbp]
@@
 
-The mentioned Non-linearities of ECG signals like powerline interference will not impact the classification result while using the wavelet transform for feature extraction because when we select the features obtained from wavelet decomposition, we would ignore the data present in trivial frequency bands (containing non-essential data/ noise) and we would not feed that non-essential data to the classifier, thus denoising can be done at the feature extraction step itself.
+The mentioned Non-linearities of ECG signals like powerline interference will not impact the classification result while using the wavelet transform for feature extraction because when we select the features obtained from wavelet decomposition, we would ignore the data present in trivial frequency bands (containing non-essential data/ noise) and we would not feed that non-essential data to the classifier, thus denoising can be done at the feature extraction step itself.\\
 Once we are done with the basic data pre-processing like baseline wandering removal, we need to compute the DWT/WPD coefficients for all scales.
@@
 The classifier then takes these parameters as input from every such scale which was selected by us and runs the respective classification algorithm.
-We then compare the obtained results from the ones obtained from the traditional classification methods.
-In WPD, for feature extraction purposes, 333 had a good idea about using all the coefficients of the last scale of decomposition.
+We then compare the obtained results from the ones obtained from the traditional classification methods.\\
+In WPD, for feature extraction purposes, \cite{Alpher01} had a good idea about using all the coefficients of the last scale of decomposition.
 Using this idea as the base, we checked the energy percentage distribution across coefficients and found that most of the energy was concentrated around 1-16 indexes.
-We have used these distribution parameters as our feature extraction pivots and have classified the healthy and unhealthy data sets.
+We have used these distribution parameters as our feature extraction pivots and have classified the healthy and unhealthy data sets.\\
 We will perform a comparative study for determining which type of mother wavelet is best suited for this application and up to what scale (levels) of cascading DWT/WPD is useful for this process.
@@
 So in order to reduce the computationally convenient and add more data, we chose to take individual sections of the ECG that had problems as our unhealthy signal and portions of ECG signal that did not have any arrhythmia as healthy ones.
-Each of the unhealthy and healthy signal length lasted for a minute.
+Each of the unhealthy and healthy signal lengths lasted for a minute.
 We then used the input signal after doing pre-processing to get the features that we have mentioned above and fed it into the classifiers to see the performance.
@@
 Scales of 9 and 10 were used for DWT and scales of 5 and 9 were used for WPD.
-The Table 1 and Table 2 shows us the accuracy results of the classification for a particular type of classifier.
+Table 1 and Table 2 show us the accuracy results of the classification for a particular type of classifier.
 
@@
 
-As we can see from the results, in general the classification accuracy of using WPD feature extraction is better than using DWT even at lower scales.
+As we can see from the results, in general, the classification accuracy of using WPD feature extraction is better than using DWT even at lower scales.
 This result was expected because of the fact that WPD will have more information when compared to the DWT of the same scale.
 \\
-As observed in Table 1 , when using DWT as feature extraction method we have to use scale up to 9 \& 10 for getting a reasonable classification result.
+As observed in Table 1, when using DWT as a feature extraction method we have to use a scale up to 9 \& 10 for getting a reasonable classification result.
 When we chose mother wavelets like Haar and Symlets, we see the classification accuracy is not as good as Daubechies mother wavelets.
-Specifically db2 provides the best accuracy at both scales 9 and 10, with accuracy reaching 86\% at scale 10.
-Now, observing the accuracy results improvement across scales, we see that there is a decent increment in accuracy when we using higher scales like scales 8-10, but if we further increase the number of levels, the classification accuracy would start saturating due to the non-availability of meaningful samples as it is too down-sampled.
-
+Specifically, db2 provides the best accuracy at both scales 9 and 10, with accuracy reaching 86\% at scale 10.
+Now, observing the accuracy results improvement across scales, we see that there is a decent increment in the accuracy when we use higher scales like scales 8-10, but if we further increase the number of levels, the classification accuracy would start saturating due to the non-availability of meaningful samples as it is too down-sampled.
 %% --------------%%
@@
 In Table 2, we see classification accuracy results while using WPD as the feature extraction method.
-Here, for scales comparison we are using scale 5 and scale 9 because we found that even scale 5 was able to match the classification accuracy of DWT.
-We tried to analyse the accuracy increment across higher scales.
+Here, for scales comparison, we are using scale 5 and scale 9 because we found that even scale 5 was able to match the classification accuracy of DWT.
+We tried to analyze the accuracy increment across higher scales.
 So we decided to compare scale 9 which was used in DWT to have a credible benchmark for our comparison.
-As we can interpret from this table the Haar and Symlets mother wavelets results are similar irrespective of the feature extraction method, however we see that the performance of db6 yields the best results when using WPD and almost matches with db2's performance in DWT even when using just scale 5 in WPD.\\
+As we can interpret from this table the Haar and Symlets mother wavelets results are similar irrespective of the feature extraction method, however, we see that the performance of db6 yields the best results when using WPD and almost matches with db2's performance in DWT even when using just scale 5 in WPD.\\
 
@@
 %% --------------------------%%
- Overall, we can say that db6 mother wavelet performs the best for our application closely followed by db2.
-Also, for our particular application, when using DWT feature extraction method, we have to use higher scales like scales 9-10 to achieve decent accuracy.
-While we would only need minimum decomposition scale 5 to achieve same accuracy in WPD method.
-However, we can further increase the accuracy of WPD method by increasing the decomposition levels up to 9-10.
+Overall, we can say that the db6 mother wavelet performs the best for our application closely followed by db2.
+Also, for our particular application, when using the DWT feature extraction method, we have to use higher scales like scales 9-10 to achieve decent accuracy.
+While we would only need a minimum decomposition scale of 5 to achieve the same accuracy in the WPD method.
+However, we can further increase the accuracy of the WPD method by increasing the decomposition levels up to 9-10.
 %%----------------------------%%
@@
 \hline
-SVM Classifier                                                   & scale 9  & scale10  & \multicolumn{2}{c|}{}                                                                & scale 5  & scale9   \\ \hline
+SVM Classifier                                                   & Scale 9  & Scale10  & \multicolumn{2}{c|}{}                                                                & Scale 5  & Scale9   \\ \hline
 DWT                                                              & Accuracy & Accuracy & \multicolumn{2}{c|}{WPD}                                                             & Accuracy & Accuracy \\ \hline
@@
 \scalebox{0.7}{
-\begin{tabular}{|l|l|l|ll|l|l|}
+\begin{tabular}{|c|c|c|c|c|c|}
 \hline
-\begin{tabular}[c]{@{}l@{}}Decision Tree \\ Classifier\end{tabular} & scale 9  & scale10  & \multicolumn{2}{l|}{}                                                                 & scale 5  & scale9   \\ \hline
-DWT                                                                 & Accuracy & Accuracy & \multicolumn{2}{l|}{WPD}                                                              & Accuracy & Accuracy \\ \hline
-\begin{tabular}[c]{@{}l@{}}Mother \\ Wavelet \\ Db6\end{tabular}    & 72       & 84       & \multicolumn{2}{l|}{\begin{tabular}[c]{@{}l@{}}Mother \\ Wavelet \\ Db6\end{tabular}} & 82       & 83       \\ \hline
-\end{tabular}}
+\begin{tabular}[c]{@{}c@{}}Decision Tree\\ Classifier\end{tabular} & Scale 9  & Scale 10 &                                                                  & Scale 5  & Scale 9  \\ \hline
+DWT                                                                & Accuracy & Accuracy & WPD                                                              & Accuracy & Accuracy \\ \hline
+\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet \\ Db6\end{tabular}   & 72       & 84       & \begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet \\ Db6\end{tabular} & 82       & 83       \\ \hline
+\end{tabular}
+}
 \caption{Decision Tree - Classification Accuracy :- \\ DWT vs WPD}
@@
 %%------------------------------%%
-We also need to check these methods efficacy with respect to different classifiers like decision tree and Support Vector Machine (SVM), K- Nearest Neighbours (KNN).
+We also need to check these methods' efficacy with respect to different classifiers like a decision tree and Support Vector Machine (SVM), K- Nearest Neighbours (KNN).
 So we decided to take the best mother wavelet (in our case db6) and optimal decomposition scales of 5 \& 9 for WPD and 9 \& 10 for DWT respectively.
-since we had already used KNN as the base classifier for table 1 \& 2 , hence we decided to compare the accuracy results using SVM and Decision Tree.\\
-As we can see from Tables 3 \& 4, the decision tree classifiers performs the best of for scale 10 DWT feature extraction.
+since we had already used KNN as the base classifier for table 1 \& 2, hence we decided to compare the accuracy results using SVM and Decision Tree.\\
+As we can see from Tables 3 \& 4, the decision tree classifiers perform the best of for scale 10 DWT feature extraction.
 whereas, with regards to scale 5, the WPD's accuracy results are equivalent to DWT's best results.
 For SVM Classifiers, WPDs performance across scales is superior to DWT's performance.\\
-As a result, we can say that when we are using KNN or SVM, WPD feature extraction is superior and is decomposition level agnostic, whereas in decision tree's case for higher levels of decomposition one can opt for DWT method of decomposition.\\
+As a result, we can say that when we are using KNN or SVM, WPD feature extraction is superior and is decomposition level agnostic, whereas in the decision tree's case for higher levels of decomposition one can opt for the DWT method of decomposition.\\
 As we can infer from the above discussion, using wavelets can simplify our task of classification of ECG signals to a great extent.
 However, the method also comes with an additional cost of Wavelet Packet Transform time overhead.
-Also, for non-binary classification purposes we need purely adaptive method of scale and mother wavelet selection which is out of the scope of this project.
+Also, for non-binary classification purposes, we need a purely adaptive method of scale and mother wavelet selection which is out of the scope of this project.
 
@@
 Using wavelet decomposition also doesn’t negatively affect the amount of useful information sent to the classifier that is used for classification.
-We were also able to improve the accuracy of the classification results to a good extent. We con
-The filtering step in the pre-processing stage is being done at the feature extraction step itself instead of using separate stages.
-This resulted in the system being more time-efficient and the complexity of the overall system was hence reduced.
+We were also able to improve the accuracy of the classification results to a good extent. The resulting system is more time-efficient and the complexity of the overall system was hence reduced.
 Further, we can say that this method can be easily applied in multi-class classification for detecting different types of abnormalities like arrhythmia, atrial flutter, fibrillation, etc to make the diagnosis more precise.

### 2021-12-12 PaperForReview.tex
final commit

@@
 \label{sec:intro}
-In today’s age Rising Cardiovascular diseases, pegs need of timely diagnosis.
+In today’s age rising Cardiovascular diseases, pegs need of timely diagnosis.
 Apart from regular health checkups, an easy \& accessible heart health diagnosis solution is needed to prevent mishaps like heart failure.
@@
 These noises can be removed by utilizing a low pass notch filter at 50/60Hz usually.
- But [Cite 1] also suggests further developments that have been used to remove the noise more efficiently.
+ But \cite{Alpher05} also suggests further developments that have been used to remove the noise more efficiently.
 Once the pre-processing is done, different ways are used to extract features which are then fed into the classifier.
@@
 Except for the Fourier transform technique, other techniques mainly rely on time-based abnormality detections.
- In the Peak detection technique, the distance between the peaks acts as the main classification basis which is computed as the distance between two R points in the P-QRS-T wave [cite 2].
+ In the Peak detection technique, the distance between the peaks acts as the main classification basis which is computed as the distance between two R points in the P-QRS-T wave.
 A simple method of classification is the Principle component analysis of the signal itself though which the classifier will handle by itself.
@@
 \subsection{Simulation Environment}
-The simulation environment consisted of us extracting the data from [repo link] which consist of patients with various heart arrhythmia.
+The simulation environment consisted of us extracting the data from the \href{https://physionet.org/content/mitdb/1.0.0/}{database} which consist of patients with various heart arrhythmia.
 Each data was of length 30 mins.
@@
 \begin{tabular}[c]{@{}c@{}}KNN \\ Classifier\end{tabular} & \multicolumn{2}{c|}{Scale : 9}                                                            & \multicolumn{2}{c|}{Scale : 10}                                                           \\ \hline
-Method                                                    & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\end{tabular}} & Accuracy & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\end{tabular}} & Accuracy \\ \hline
+Method                                                    & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\end{tabular}} & Accuracy\% & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\end{tabular}} & Accuracy\% \\ \hline
 \multirow{5}{*}{DWT}                                      & \multicolumn{1}{c|}{Db2}                                                       & 80       & \multicolumn{1}{c|}{Db2}                                                       & 86       \\ \cline{2-5}
@@
 \begin{tabular}[c]{@{}c@{}}KNN \\ Classifier\end{tabular} & \multicolumn{2}{c|}{Scale : 5}                                                            & \multicolumn{2}{c|}{Scale : 9}                                                            \\ \hline
-Method                                                    & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\end{tabular}} & Accuracy & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\end{tabular}} & Accuracy \\ \hline
+Method                                                    & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\end{tabular}} & Accuracy\% & \multicolumn{1}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\end{tabular}} & Accuracy\% \\ \hline
 \multirow{5}{*}{WPD}                                      & \multicolumn{1}{c|}{Db2}                                                       & 80       & \multicolumn{1}{c|}{Db2}                                                       & 86       \\ \cline{2-5}
@@
 SVM Classifier                                                   & Scale 9  & Scale10  & \multicolumn{2}{c|}{}                                                                & Scale 5  & Scale9   \\ \hline
-DWT                                                              & Accuracy & Accuracy & \multicolumn{2}{c|}{WPD}                                                             & Accuracy & Accuracy \\ \hline
+DWT                                                              & Accuracy\% & Accuracy\% & \multicolumn{2}{c|}{WPD}                                                             & Accuracy\% & Accuracy\% \\ \hline
 \begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet \\ Db6\end{tabular} & 81.4     & 85.3     & \multicolumn{2}{c|}{\begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet\\ Db6\end{tabular}} & 83       & 87       \\ \hline
@@
 \begin{tabular}[c]{@{}c@{}}Decision Tree\\ Classifier\end{tabular} & Scale 9  & Scale 10 &                                                                  & Scale 5  & Scale 9  \\ \hline
-DWT                                                                & Accuracy & Accuracy & WPD                                                              & Accuracy & Accuracy \\ \hline
+DWT                                                                & Accuracy\% & Accuracy\% & WPD                                                              & Accuracy\% & Accuracy\% \\ \hline
 \begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet \\ Db6\end{tabular}   & 72       & 84       & \begin{tabular}[c]{@{}c@{}}Mother \\ Wavelet \\ Db6\end{tabular} & 82       & 83       \\ \hline
@@
 Using wavelet decomposition also doesn’t negatively affect the amount of useful information sent to the classifier that is used for classification.
-We were also able to improve the accuracy of the classification results to a good extent. The resulting system is more time-efficient and the complexity of the overall system was hence reduced.
+We were also able to improve the accuracy of the classification results to a good extent.
 Further, we can say that this method can be easily applied in multi-class classification for detecting different types of abnormalities like arrhythmia, atrial flutter, fibrillation, etc to make the diagnosis more precise.
